---
title: "PS270 Final Project"
author: "Pete Xu"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Project thoughts

I plan to investigate whether higher levels of economic development are associated with greater life satisfaction across countries. My research question is: Does GDP per capita predict average life satisfaction? Drawing on modernization theory, I hypothesize that countries with higher GDP per capita will report higher levels of life satisfaction because greater material wealth improves living standards, economic security, and access to public goods, which in turn shape individual well-being in systematic ways. I will use data from the World Happiness Report. The dependent variable is the Life Ladder score, which measures average self-reported life satisfaction on a scale from 0 to 10 for each country-year. The key explanatory variable is GDP per capita (logged), measured in constant U.S. dollars. Both variables exhibit substantial variation across countries and over time, making regression analysis appropriate. If my hypothesis is correct, I expect to observe a positive and statistically significant relationship between GDP per capita and life satisfaction. A null or negative coefficient would disconfirm the hypothesis.













```{r}
library(tidyverse)

happiness <- read_csv("data/World_Happiness_Report.csv", show_col_types = FALSE)

head(happiness)

happiness2015 <- read_csv("data/2015.csv",show_col_types = FALSE)
head(happiness2015)
```
```{r}
library(tidyverse)

happiness_a <- read_csv("data/World_Happiness_Report.csv", show_col_types = FALSE)
happiness_b <- read_csv("data/2015.csv", show_col_types = FALSE)

head(happiness_a)
head(happiness_b)
```



```{r}


happiness_a_clean <- happiness_a %>%
  select(Country, `Happiness Score`, Economy, Family, Health, Freedom, Generosity, Corruption) %>%
  drop_na()
happiness_b_clean <- happiness_b %>%
  select(Country, Region, `Happiness Score`, 
         `Economy (GDP per Capita)`, Family, 
         `Health (Life Expectancy)`, Freedom, Generosity, 
         `Trust (Government Corruption)`) %>%
  drop_na()


head(happiness_a_clean)
head(happiness_b_clean)
```

```{r}
ggplot(happiness_a_clean, aes(x = Economy, y = `Happiness Score`)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1) +
  labs(
    title = "Economic Development and Happiness",
    subtitle = "World_Happiness_Report.csv",
    x = "Economy",
    y = "Happiness Score"
  )
```
```{r}
ggplot(happiness_b_clean, aes(x = `Economy (GDP per Capita)`, y = `Happiness Score`)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1) +
  labs(
    title = "GDP per Capita and Happiness",
    subtitle = "2015.csv",
    x = "Economy (GDP per Capita)",
    y = "Happiness Score"
  )
```
```{r}
model_b <- lm(`Happiness Score` ~ `Economy (GDP per Capita)`, data = happiness_b_clean)

summary(model_b)
```
```{r}
library(broom)

model_b %>%
  tidy() %>%
  knitr::kable(
    digits = 3,
    caption = "Regression Results: GDP per Capita and Happiness Score"
  )
```
To answer my research question, I ran a regression with Happiness Score as the dependent variable and GDP per capita as the independent variable. The results show a positive relationship between GDP per capita and happiness. The estimated coefficient of 2.218 indicates that increases in GDP per capita are associated with substantial increases in happiness scores, reinforcing the strength of this positive relationship. In other words, when GDP per capita increases by 1 unit, the happiness score increases by about 2.22 points. This means that countries with higher GDP per capita tend to have higher happiness scores. 

